---
title: "Combine data frames with new column"
author: "Jamie Hudson, Data Scientist"
date: "2022-06-10"
categories: ["dplyr", "gdata"]
tags: ["cleaning", "plot", "tidyverse"]
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### *Answer originally provided by Jack Penzer & Jamie Hudson, Data Scientists*

## The Problem

Sometimes we have multiple data frames containing the same variables but for different companies, for example. We might then want to be able to join these data frames together, creating a new variable (column) that corresponds to the name of the original data frame.

In this example, we are going to use three dummy data frames called `amazon`, `microsoft`, and `google`, which contain the columns `message_id`, `message`, and `author`- though the solutions provided would work even if the data frames had many more columns.

```{r, eval = TRUE, echo = FALSE, message = FALSE}
library(tidyverse)
amazon <- tribble(
  ~message_id, ~message, ~author,
  "12345", "Amazon was founded in 1994 from my Garage in Bellevue, Washing", "Jeff Bezos",
  "23456", "What do you call two monkeys that share an Amazon Account? Prime mates", "Tom Smith",
  "34567", "30% of the world's species live in the Amazon", "David Attenborough"
)

microsoft <- tribble(
  ~message_id, ~message, ~author,
  "123", "The name Microsoft is a combination of the words microcomputer and software", "Mike O'Soft",
  "456", "MS Office was released on the Macintosh before the Windows OS", "@facts_r_us",
  "789", "RT @facts_r_us MS Office was released on the Macintosh before the Windows OS", "@retweet_bot"
)

google <- tribble(
  ~message_id, ~message, ~author,
  "22", "A Googol is a 1 with 100 zeros after it", "Larry Page",
  "33", "Check out these amazing Google facts that will blow your mind bit.ly", "Elgoog",
  "44", "Google is the world's most visited website", "Google Fan"
)
```

```{r}
head(amazon)
head(microsoft)
head(google)
```

## Solution 1- `mutate()` and then `bind_rows()`

One solution is to add a new variable to each data frame individually using the `dplyr::mutate()` function, and then bind these individual data frames into one using the `dplyr::bind_rows()` function.

```{r}
library(dplyr)

amazon_mutated <- amazon %>% 
  dplyr::mutate(source = "amazon")

microsoft_mutated <- microsoft %>% 
  dplyr::mutate(source = "microsoft")

google_mutated <- google %>% 
  dplyr::mutate(source = "google")

combined_df <- dplyr::bind_rows(amazon_mutated, microsoft_mutated, google_mutated)
```

Here, we used `dplyr::mutate()` to create a new column called `source`, which we labelled with the name of each data frame, before binding using `dplyr::bind_rows()` to produce our new data frame called `combined_df`. We can view what this data frame looks like, with our new `source` column:

```{r}
head(combined_df)
```

## Solution 2- quicker with `gdata::combine()`

A quicker and simpler way to achieve the same result is using the `gdata::combine()` function. This combines our data frames into rows of a new data frame with the additional column `source`, all at once. The value of this new `source` column is the name of the original (source) data frame. 

**N.B. the first time you run this function, you may have to install the `gdata` library**

```{r, eval=FALSE}
install.packages("gdata")
```


```{r, message = FALSE}
library(gdata)

combined_df <- gdata::combine(amazon, microsoft, google)

head(combined_df)
```

Et voilà, now we have a new data frame with an additional `source` column with just one line of code! 
