---
title: "Change text size of SegmentR plots"
author: "Jamie Hudson, Data Scientist"
date: "2022-06-29"
categories: ["tidyverse", "ggplot2", "segmentr"]
tags: ["ggplot2", "ggsave", "figures"]
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### *Answer originally suggested by Jack Penzer, Senior Data Scientist*

## The Problem

Sometimes, the plots produced by packages such as SegmentR appear squished with not much white space surrounding the axis text (i.e. the terms for each topic).

Let's use the example 'Top Terms' plot from the [SegmentR documentation](https://avery-island.github.io/SegmentR/articles/1_topic_modelling.html) to see what this means. We start by quickly running through a very simple topic modelling workflow (NB we are not trying different values of `min_freq`, `k_opts` etc here because this is just for demonstration- we **must** try different values for project work)

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(SegmentR)
library(JPackage)
```


```{r message=FALSE}
sprinklr_export <- SegmentR::sprinklr_export %>%
  filter(!str_detect(Message, "A word of advice\\?"))

dtms <- sprinklr_export %>%
  SegmentR::make_DTMs(text_var = Message,
                      min_freq = 10,
                      hashtags = FALSE,
                      mentions = FALSE,
                      clean_text = TRUE,
                      remove_stops = TRUE)

ldas <- dtms %>%
  SegmentR::fit_LDAs(k_opts = 3,
                     iter_opts = 1000)

explore <- ldas %>%
  SegmentR::explore_LDAs(top_terms = TRUE,
                         diff_terms = TRUE,
                         bigrams = TRUE,
                         exemplars = TRUE)
```

Now, we can view this plot by `pull`ing the top_terms and `pluck`ing "all_terms" (again, this is all explained in the [SegmentR documentation](https://avery-island.github.io/SegmentR/articles/1_topic_modelling.html))

```{r}
explore %>%
  filter(freq_cutoff == 10 & k == 3 & iter == 1000) %>%
  pull(top_terms) %>%
  pluck(1, "all_terms")
```

We can see that all of the text on the y axis (the top terms for each topic) is a tad squished and doesn't look the best.

## The Solution

There are a few different solutions which can be tried to fix this problem. The first is to save the plot using `JPackage::plot_save()` with larger dimensions. Please read the Slack Overflow page for this [here](https://slack-overflow-help.netlify.app/post/2023-01-13-ggsave/ggsave/).

If this doesn't work, we can manipulate the plot by either :
  1) reducing the number of terms shown, 
  2) laying out the plots on only one row,
  3) or reducing the font size

Let's look at these solutions one by one.

1) To reduce the number of terms shown, we need to go back to our `SegmentR::explore_LDAs()` function. There is an argument here, `top_n`, which determines how many terms per topic we visualise. For a full list of potential arguments, you can type `?SegmentR::explore_LDAs` to access and read the documentation. Let's change this argument to 15 and see if this helps (note as seen in the documentation, the default value for this is 20):

```{r}
explore <- ldas %>%
  SegmentR::explore_LDAs(top_terms = TRUE,
                         diff_terms = TRUE,
                         bigrams = TRUE,
                         exemplars = TRUE,
                         top_n = 15)

explore %>%
  filter(freq_cutoff == 10 & k == 3 & iter == 1000) %>%
  pull(top_terms) %>%
  pluck(1, "all_terms")
```

Et voilá, you can see that now we have reduced the number of terms, there is nice white space between the bars and the text isn't so squished. 

2) We can also use the `SegmentR::explore_LDAs` to force the plot to appear in one row (rather than facetting the plot into two rows of charts, as it is currently doing). Again, this is controlled by the argument `nrow` and we can assign this to be equal to `1`:

```{r}
explore <- ldas %>%
  SegmentR::explore_LDAs(top_terms = TRUE,
                         diff_terms = TRUE,
                         bigrams = TRUE,
                         exemplars = TRUE,
                         nrow = 1)

explore %>%
  filter(freq_cutoff == 10 & k == 3 & iter == 1000) %>%
  pull(top_terms) %>%
  pluck(1, "all_terms")
```

As we are not forcing the plots on two rows, the stretch out vertically a bit more now.

3) Finally, we can edit the size of the text specifically. In contrast to the previous two solutions, we do not do this via the `SegmentR::explore_LDAs` function, but rather explicitaly set this when pulling the plot from the `explore` object. We do this by providing the `ggplot2` function `theme()` to our code. Within `theme()` we can specify the argument `axis.text.y` to be equal to `element_text(size = 8)`. Play around with this size (larger numbers = larger text) until you are happy. Note that this can also be used to specify the size of the x axis text with the argument `axis.text.y`. As always though, if you end up changing the size of the axis text in one plot, be considerate as to whether other plots also need to be changed for consistency. 


```{r}
explore <- ldas %>%
  SegmentR::explore_LDAs(top_terms = TRUE,
                         diff_terms = TRUE,
                         bigrams = TRUE,
                         exemplars = TRUE)

explore %>%
  filter(freq_cutoff == 10 & k == 3 & iter == 1000) %>%
  pull(top_terms) %>%
  pluck(1, "all_terms") +
  theme(axis.text.y = element_text(size = 8))
```

And there you go, there are some potential solutions to helping clean up some plots that have been made in SegmentR.
